#Assignment-1
##Loading the data
```{r}
activity <- read.csv("activity.csv", colClasses = c("numeric", "character", 
    "numeric"))
names(activity)
library(lattice)
activity$date <- as.Date(activity$date, "%Y-%m-%d")
```
#What is mean total number of steps taken per day?
```{r}
StepsTotal <- aggregate(steps ~ date, data = activity, sum, na.rm = TRUE)
```
#The histogram is
```{r, echo=FALSE}
hist(StepsTotal$steps, main = "Total steps by day", xlab = "day", col = "red")
```
#And the mean and median  is
```{r}
mean(StepsTotal$steps)
## [1] 10766
median(StepsTotal$steps)
## [1] 10765
```
#Make a data frame first with the values needed. Create a data frame with the days and the total of steps by day
```{r}
steps <- rep(NA, 61)
day <- rep("NA", 61)
stepsday <- tapply(activity$steps, activity$date, sum, na.rm = T)
length(stepsday)
## [1] 61
for (i in 1:61) {
    steps[i] <- stepsday[[i]]
    day[i] <- names(stepsday)[i]
}

df <- data.frame(day, steps)
head(df)
##          day steps
## 1 2012-10-01     0
## 2 2012-10-02   126
## 3 2012-10-03 11352
## 4 2012-10-04 12116
## 5 2012-10-05 13294
## 6 2012-10-06 15420
```
#The histogram
```{r}
mean(StepsTotal$steps)
## [1] 10766
median(StepsTotal$steps)
## [1] 10765
```
#Make a data frame first with the values needed. Create a data frame with the days and the total of steps by day
```{r}
```{r}
mean(StepsTotal$steps)
## [1] 10766
median(StepsTotal$steps)
## [1] 10765
```
#Make a data frame first with the values needed. Create a data frame with the days and the total of steps by day
```{r}
steps <- rep(NA, 61)
day <- rep("NA", 61)
stepsday <- tapply(activity$steps, activity$date, sum, na.rm = T)
length(stepsday)
## [1] 61
for (i in 1:61) {
    steps[i] <- stepsday[[i]]
    day[i] <- names(stepsday)[i]
}

df <- data.frame(day, steps)
head(df)
##          day steps
## 1 2012-10-01     0
## 2 2012-10-02   126
## 3 2012-10-03 11352
## 4 2012-10-04 12116
## 5 2012-10-05 13294
## 6 2012-10-06 15420
```
#The histogram
```{r}
hist(df$steps, main = "Total steps by day", xlab = "day", col = "green")
```
## Average daily activity pattern
#Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
time_series <- tapply(activity$steps, activity$interval, mean, na.rm = TRUE)
```
#Plot is

```{r}
plot(row.names(time_series), time_series, type = "l", xlab = "5-min interval", 
    ylab = "Average across all Days", main = "Average number of steps taken", 
    col = "red")
```
#5min interval, on avearage across all the days. 
```{r}
max_interval <- which.max(time_series)
names(max_interval)
## [1] "835"
```
### Inputting missing values
#Calcutale teh total number of missing values (number of rows with NAs)
```{r}
activity_NA <- sum(is.na(activity))
activity_NA
## [1] 2304
```
#use the mean/median for that day, or the mean for that 5-minute interval and replace  mean in 5 min interval
```{r}
StepsAverage <- aggregate(steps ~ interval, data = activity, FUN = mean)
fillNA <- numeric()
for (i in 1:nrow(activity)) {
    obs <- activity[i, ]
    if (is.na(obs$steps)) {
        steps <- subset(StepsAverage, interval == obs$interval)$steps
    } else {
        steps <- obs$steps
    }
    fillNA <- c(fillNA, steps)
}
```
#New data set with missing data filled in
```{r}
new_activity <- activity
new_activity$steps <- fillNA
```
##Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?
```{r}
StepsTotal2 <- aggregate(steps ~ date, data = new_activity, sum, na.rm = TRUE)
```
##Histogram
```{r}
hist(StepsTotal2$steps, main = "Total steps by day", xlab = "day", col = "red")
```
#And the mean and median is
```{r}
mean(StepsTotal2$steps)
## [1] 10766
median(StepsTotal2$steps)
## [1] 10766
```
###Are there differences in activity patterns between weekdays and weekends?
#Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
```{r}
day <- weekdays(activity$date)
daylevel <- vector()
for (i in 1:nrow(activity)) {
    if (day[i] == "Saturday") {
        daylevel[i] <- "Weekend"
    } else if (day[i] == "Sunday") {
        daylevel[i] <- "Weekend"
    } else {
        daylevel[i] <- "Weekday"
    }
}
activity$daylevel <- daylevel
activity$daylevel <- factor(activity$daylevel)

stepsByDay <- aggregate(steps ~ interval + daylevel, data = activity, mean)
names(stepsByDay) <- c("interval", "daylevel", "steps")
```
#Plot
```{r}
xyplot(steps ~ interval | daylevel, stepsByDay, type = "l", layout = c(1, 2), 
    xlab = "Interval", ylab = "Number of steps")
```
